### 2023-09-04 sections/3-datasets.tex
more tweaks
submitted paper

The specifications for the machines used in benchmarking are as follows:

\begin{itemize}
\item Comparisons to FAISS, Bruteforce28-core Intel Xeon E5-2690 v4 2.60GHz, 512GB RAM, Manjaro Linux with kernel 5.15.78-1-MANJARO \#1 SMP PREEMPT. Rust 1.70, Python 3.10.5
\item Amazon AWS: 64-core Intel Xeon Platinum 8375C CPU 2.90GHz, 512GB RAM, Ubuntu 22.04.3-Ubuntu SMP. Rust 1.71.1, Python 3.10.12
\end{itemize}

>>>>

All benchmarks were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}
The OS kernel was Ubuntu 22.04.3-Ubuntu SMP. 
The Rust compiler was Rust 1.71.1, and the Python interpreter version was 3.10.12.

### 2023-09-04 sections/4-results.tex
submitted paper

We report the highest throughput value for each algorithm at a recall value of 1.0. We report the recall value for each algorithm at the highest throughput value. We report the mean and median speedup factor of CAKES over each algorithm.

For $k= 10$, we report a mean $989\times$ (median $248\times$) speedup over faiss-ivf, a mean $1456\times$ (median $582\times$) speedup over bruteforce-blas, 
and a mean $63\times$ (median $49\times$) speedup over HNSW. For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.

@@

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
    \label{table:results:ann-10}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
    \cline{2-9}
    &                    Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
    \hline
    fashion-mnist      & 1.000 & 188.03                           & 1.000 & 52.91                                  & 1.000* & 339.74                                                    & 1.000 & 16,220 \\
    \hline
    gist                   & 1.000 & 3.55                           & 1.000 & 2.63                                     & 0.953 & 149.58                                                   & 1.000 & 10,600 \\
    \hline
    glove-25              & 1.000 & 78.78                          & 1.000 & 33.52                              & 1.000 & 552.67                                                  & 0.798 & 19,520 \\
    \hline
    glove-100             & 1.000 & 21.16                          & 1.000 & 16.52                                & 0.998 & 155.90                                                  & 0.904 & 6,935 \\
    \hline
    sift                  & 1.000 &  24.52                          & 1.000 & 16.40                               & 1.000 & 275.07                                                    & 1.000 & 31,800 \\
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$. Other methods did not report results for glove-25. Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
    \label{table:results:ann-100}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
    &                    Recall & QPS                           & Recall & QPS                           & Recall & QPS                                           & Recall & QPS \\
    \hline
    fashion-mnist         & 1.000 & 111.47                           & 1.000* & 16.02                                  & 1.000* & 277.80                                                    & 1.000 & 1,549 \\ 
    \hline
    gist                   & 1.000 & 2.70                           & 1.000* & 0.80                                 & 0.995 & 59.71                                                    & 1.000 & 753.9 \\
    \hline
    glove-25              & -- & --                                & -- & --                                & -- & --                                                    & 0.851 & 1,471 \\
    \hline
    glove-100             & 1.000* &  19.04                          & 1.000* & 7.52                                 & 0.987 & 92.58                                                    & 0.905 & 396.0 \\
    \hline
    sift                  & 1.000 &  23.50                          & 1.000 & 6.51                                  & 1.000* & 179.14                                                    & 1.000 & 2,852 \\                                                 
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

>>>>

We report the highest throughput value for each algorithm at a recall value of 1.0. We report the recall value for each algorithm at the highest throughput value. We report the mean and median speedup factor of CAKES over each algorithm.
For $k= 10$, we report a mean $7654\times$ (median $1200\times$) speedup over faiss-ivf, a mean $11325\times$ (median $2821\times$) speedup over bruteforce-blas, 
and a mean $485\times$ (median $417\times$) speedup over HNSW. 
For $k=100$, we report a mean $4479\times$ (median $3456\times$) speedup over faiss-ivf, a mean $16010\times$ (median $13548\times$) speedup over bruteforce-blas, and a mean $403\times$ (median $372\times$) speedup over HNSW.

@@

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=10$}
    \label{table:results:ann-10}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
    \cline{2-9}
    &               Recall & QPS      & Recall & QPS    & Recall & QPS    & Recall & QPS \\
    \hline
    fashion-mnist & 1.000  & 188.03    & 1.000 & 52.91  & 1.000* & 339.74 & 1.000 & 141,800 \\
    \hline
    gist          & 1.000 & 3.55  & 1.000 & 2.63        & 0.953 & 149.58  & 1.000 & 76,420 \\
    \hline
    glove-25      & 1.000 & 78.78  & 1.000 & 33.52      & 1.000 & 552.67  & 0.798 & 94,580 \\
    \hline
    glove-100     & 1.000 & 21.16  & 1.000 & 16.52      & 0.998 & 155.90  & 0.904 & 4,566 \\
    \hline
    sift          & 1.000 &  24.52 & 1.000 & 16.40      & 1.000 & 275.07  & 1.000 & 357,400 \\
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$. Other methods did not report results for glove-25. Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
    \label{table:results:ann-100}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{faiss-ivf}} & \multicolumn{2}{|c|}{\textbf{bruteforce-blas}} & \multicolumn{2}{|c|}{\textbf{hnsw(faiss)}} & \multicolumn{2}{|c|}{\textbf{CAKES}} \\
    &                    Recall & QPS  & Recall & QPS     & Recall & QPS      & Recall & QPS \\
    \hline
    fashion-mnist  & 1.000 & 111.47    & 1.000* & 16.02   & 1.000* & 277.80   & 1.000 & 71,370 \\ 
    \hline
    gist           & 1.000 & 2.70      & 1.000* & 0.80    & 0.995 & 59.71     & 1.000 & 29,090 \\
    \hline
    glove-25       & -- & --           & --     & --      & -- & --           & 0.851 & 49,350 \\
    \hline
    glove-100      & 1.000* &  19.04   & 1.000* & 7.52    & 0.987 & 92.58     & 0.905 & 4,384 \\
    \hline
    sift           & 1.000 &  23.50    & 1.000  & 6.51    & 1.000* & 179.14   & 1.000 & 147,400 \\                                                 
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

### 2023-09-04 sections/5-discussion.tex
submitted paper

In future work, we plan to investigate compression by representing differences at each level of the
binary tree, particularly for string or genomic data, where all differences are discrete.
When we descend a level in the cluster tree, it takes fewer bits to encode each point in terms of the center. 
However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
This presents the following tradeoff: \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.

We also plan to use CAKES with anomaly detection. Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses 
a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of 
distance to the $k$-th nearest neighbor as an additional method in our ensemble.

Finally, we would like to add the ability to view live-updates as new points are added to a dataset. For example, 
we could build the cluster tree with a subsample of the full data set and then add the rest of the instances to 
simulate live updates.

>>>>

In future work, we plan to investigate compression by representing differences at each level of the
binary tree, particularly for string or genomic data, where all differences are discrete.
When we descend a level in the cluster tree, it takes fewer bits to encode each point in terms of the center. 
However, each descent of a level in the cluster tree increases the bit-cost of specifying a particular cluster. 
This presents the following tradeoff: \emph{How do we algorithmically determine the depth to which we should cluster a dataset to achieve optimal compression?} 
In our future work, we will explore which properties of the dataset and of clusters can inform this optimal depth.
Along with compression, we would like to add the ability to perform ``online'' updates as new points are added to a dataset.

We also plan to use CAKES with anomaly detection. Our anomaly detection tool, CHAODA~\cite{ishaq2021clustered}, uses 
a variety of methods in an ensemble to determine an anomaly score for a point. We would like to explore the effectiveness of using
distance to the $k$-th nearest neighbor as an additional method in our ensemble.
